---
title: <center>"A linear regression package in R based on RCourseLab4" </center>
author: "Farhana chowdhury Tondra,Syeda Farha Shazmeen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEngine{knitr::render(..., 'all')}
  %\VignetteIndexEntry{A linear regression package in R based on RCourseLab4}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}{inputenc}
---

**RCourseLab4** is entitled for the students who have completed lab3. This is an advanced assigment of R. This package contains several methods of linear regression model including **vignette** manual.We are using linear algebra to create the most basic functionality in the R package. It has a R script based on RC class  where we first defined the fields and then methods. The methods which implemented for calculating linear regression model is defined later.This package contains few things :

- A Method RC class is used here named as linreg(). 
- About six methods implemented for linear regression model.
- Used Roxygen skeleton and vignette for documentation.

## linreg() Class:

The function we created here is called **linreg()** and have the two arguments formula and data. The function returns an object with of class linreg. The class is implemented as Referrence class using **setRefClass()**. Outside the class we created object **linreg_mod** for returning the results. This class accepts several arguments as lists in **fields** and **methods** classes. In fields class we defined the necessary parameters we need for our class and their type. Where in methods class we defined our required calculations.

## Calculations using ordinary least squares: 

### Regressions coefficients:  
Regressions coefficient using linear algebra
$$\hat{\beta} = (\mathbf{X}^\top \mathbf{X})^{-1}(\mathbf{X}^\top \mathbf{y})$$

### The fitted values:
Fitted values using linear algebra    
              $$ \hat{y}= X\hat{\beta} $$
              
### The residuals:
Residuals using linear algebra 
        $$\hat{e}= y - \hat{y} = X\hat{\beta} $$

### The degrees of freedom:
The degrees of freedom using linear algebra 
        $$ df = n - p $$
        where n is the number of observations and p is the number of parameters in the model

### The residual variance:
The residual variance: using linear algebra 
        $$ \hat{\sigma}^2 = \frac{\mathbf{e}^\top \mathbf{e}}{df}$$
   
### The variance of the regression coefficients:        
The variance of the regression coefficients
  $$ \hat{Var}(\hat{\beta}) = \hat{\sigma}^2({\mathbf{X}^\top \mathbf{X}})^{-1} $$

###The t-values for each coefficient:
The t-values for each coefficient
        $$ t_\beta= \frac{\hat{\beta}}{\sqrt{{Var}(\hat{\beta})}} $$

## Methods:
###linreg_mod$print()
It prints out the coefficients and coefficient names.
```{r, echo=TRUE}
knitr::opts_chunk$set(error = TRUE)

print = function() {
    
                      cat("\n","Call:","\n",
                          paste("linreg(", "formula = ", formula[2]," ", formula[1], " ", 
                                formula[3],", ", "data = ", parsedata, ")",sep = "", 
                                collapse = "\n" ),"\n","Coefficients:","\n",
                          paste(row.names(beta), sep = "  ", collapse ="  " ),"\n",
                          format(round(beta,2), justify = "centre",width = 10))
                    }
linreg_mod <- linreg$new(Petal.Length~Sepal.Width+Sepal.Length,data=iris)
linreg_mod$print()

```


###linreg_mod$plot()
## Figures of two plot are shown below.
```{r fig.width=7, fig.height=4, echo=FALSE}
plot=function() {
              
              library(ggplot2)
              library(ggThemeAssist)
              
  #Theme             
              liu_graphic_theme<-   theme(
                
                      panel.background = element_rect(fill="white"),
                      plot.margin = unit(c(1,1,1,1), "cm"),
                      plot.caption = element_text(size=12, hjust=0.5, margin=margin(t=15)),
                      plot.title = element_text(color="#666666", face="bold", size="15",hjust=0.5),
                      panel.grid.major.y = element_blank(),
                      panel.grid.minor.y = element_blank(),
                      panel.grid.major.x = element_blank(),
                      panel.grid.minor.x = element_blank(),
                      axis.line = element_line(color= "#666666", size=0.1),
                      axis.text.x = element_text(color="#666666", size="5"),
                      axis.text.y = element_text(color="#666666", size="5"),
                      axis.title.x = element_text(color="#666666", size="12", face="bold"),
                      axis.title.y = element_text(color="#666666", size="12", face="bold"),
                      axis.ticks.x = element_line(color = "blue", size = 0.3)
                       )
               
              b<-expression(bold(sqrt(bold("Standardized Residuals"))))
              caption<-  paste("linreg(", formula[2]," ", formula[1], " ", formula[3], ")")
   #Plot1           
              df<- data.frame(FittedValues=ybar,Residulas=ebar)
              Plot1<-ggplot(df, aes(x = FittedValues, y = Residulas))+
                     geom_point(shape = 21, colour = "#666666", fill = "white", size = 3, stroke = 1.5) + ggtitle("Residual vs Fitted") +
                     geom_abline(slope = 0, intercept = 0,linetype = "dotted") +
                     geom_smooth(method = "loess",color = "red", se = FALSE) +
                     ylab("Residuals") + xlab("Fitted values")+labs(caption=caption)+
                     liu_graphic_theme
            
              df1<- data.frame(FittedValues=ybar,Residulas1=stdres)
  #Plot2        
            
              Plot2<-ggplot(df1, aes(x = FittedValues, y = Residulas1))+
                     geom_point(shape = 21, colour = "#666666", fill = "white", size = 3, stroke = 1.5) + ggtitle("Scale-Location") +
                     geom_abline(slope = 0, intercept = 0,linetype = "dotted") +
                     geom_smooth(method = "loess",color = "red", se = FALSE) +ylab(b)+
                     labs(caption=caption)+
                     xlab("Fitted values")+ 
                     liu_graphic_theme
             
                     return(list("Residual vs Fitted"= Plot1,"Scale Location"= Plot2))
            
          }        

linreg_mod <- linreg$new(Petal.Length~Sepal.Width+Sepal.Length,data=iris)
linreg_mod$plot()
```

###linreg_mod$resid()
resid() returns the vector of residuals of e.
```{r, echo=TRUE}
knitr::opts_chunk$set(error = TRUE)
resid = function(){
                     cat("Vector of Residuals:","\n")
                     return(as.vector(round(ebar, 2)))
           }
linreg_mod <- linreg$new(Petal.Length~Sepal.Width+Sepal.Length,data=iris)
linreg_mod$resid()

```


###linreg_mod$pred()
pred() returns the predicted values of ybar.
```{r, echo=TRUE}
knitr::opts_chunk$set(error = TRUE)
pred =function(){
                     cat("Predicted values or fitted values:","\n")
                     return(as.vector(round(ybar, 2)))
          }
linreg_mod <- linreg$new(Petal.Length~Sepal.Width+Sepal.Length,data=iris)
linreg_mod$pred()
```

###linreg_mod$coef()
coef() returns the coefficients as a named vector

```{r, echo=TRUE}
knitr::opts_chunk$set(error = TRUE)
coef = function() {
                     cat("Regressions coefficients:","\n\n")
                     return(as.vector(round(beta,2)))
          }
linreg_mod <- linreg$new(Petal.Length~Sepal.Width+Sepal.Length,data=iris)
linreg_mod$coef()
```

###linreg_mod$summary()
summary() computes and returns a list of summary statistics of the fitted linear model given in object, using the components (list elements).

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
summary = function() {
                     output <- data.frame( rname = rownames(beta),
                          Estimate =round(beta,2),Std_err = round(sqrt(diag(var_beta)),2),t_value = round(t_beta, 2),
                          pr = round(p, 4))
                     output$rname <- as.character(output$rname)
                     row.names(output) <- NULL
                     output <- rbind(c("","Estimate", "Std. Error", "t value", "Pr(>|t|)"), output)
                     
                     for(i in 2:nrow(output)){
                     
                       if(output$pr[i] == 0){
                       output$pr[i] <- "***"
                     } else if(output$pr[i] > 0 & output$pr[i] <= 0.001){
                       output$pr[i] <- "**"
                     } else if(output$pr[i] > 0.001 & output$pr[i] <= 0.01){
                       output$pr[i] <- "*"
                     } else if(output$pr[i] > 0.01 & output$pr[i] <= 0.05){
                       output$pr[i] <- "."
                     } else if(output$pr[i] > 0.05 & output$pr[i] <= 0.1){
                       output$pr[i] <- " "
                     } else if(output$pr[i] > 0.1){
                       output$pr[i] <- " "
                     }
                   }
                    cat("\n Call:\n\n",
                    paste("linreg(", "formula = ", formula[2]," ", formula[1], " ", formula[3],", ", "data = ", parsedata, ")",sep = "", collapse = "\n" ),
                    "\n\n Coefficients:\n\n")

                    for(i in 1:ncol(output)){
                    distance <- max(nchar(as.character(output[, i])), na.rm = TRUE)
                     for(j in 1:nrow(output)){
                       output[j, i] <- format(output[j, i], width=distance,justify = c("right"))
                     }
                     }
                              
                    for(i in 1:nrow(output)){
                    cat(paste(as.character(output[i, ]),collapse = " "),"\n")
                    }
                    cat("\n Residual standard error:",varian, "on", ddf, "degrees of freedom")
        
                       }

linreg_mod <- linreg$new(Petal.Length~Sepal.Width+Sepal.Length,data=iris)
linreg_mod$summary()
```


##Roxygen Description files:
###Package documentation : **RCourseLab4.Rd**
This file includes the detail description of how package works , who worked for it and so on.

###Function documentation : **linreg-class.Rd**
This file contains the description of class file , fields, method, examples, references and return statements.

###Vignettes documentation : **lab4.Rmd**
This is about how the whole package is made of and how to read it and implement it.
 
##Install **RCourseLab4** package for future reference:
devtools::install_github("farhashazmeen/lab4")
```{r,eval = FALSE}
devtools::install_github("farhashazmeen/lab4")
```

## Limitations:
Unable to add geographical theme on ggplot and estimation of QR decomposition.


